---
title: "03 Automated Machine Learning with H20"
date: "2020-06-09"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

```{r}
# Include libraries
library(tidymodels)
library(tidyverse)
library(magrittr)
library(dplyr)
library(sjmisc)
library(magrittr)
library(haven)
library(sjlabelled)
library(stringr)

# Read the dataset
employee_attrition_tbl <- read.csv("./raw_data/datasets-1067-1925-WA_Fn-UseC_-HR-Employee-Attrition.txt")
dept_job_role_tbl <- employee_attrition_tbl %>% select(EmployeeNumber, Department, JobRole, PerformanceRating, Attrition)

dept_job_role_tbl %>%
  group_by(Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))
```
```{r}
# Attrition by department
dept_job_role_tbl %>%
  
  # Block 1
  group_by(Department, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  
  # Block 2: Caution: It's easy to inadvertently miss grouping when creating counts & percents within groups
  group_by(Department) %>%
  mutate(pct = n / sum(n))
```
```{r}
# Attrition by job role
dept_job_role_tbl %>%
  
  # Block 1
  group_by(Department, JobRole, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  
  # Block 2
  group_by(Department, JobRole) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  
  # Block 3
  filter(Attrition %in% "Yes")
```
```{r}
# Develop KPI
dept_job_role_tbl %>%
  
  # Block 1
  group_by(Department, JobRole, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  
  # Block 2
  group_by(Department, JobRole) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  
  # Block 3
  filter(Attrition %in% "Yes") %>%
  arrange(desc(pct)) %>%
  mutate(
    above_industry_avg = case_when(
      pct > 0.088 ~ "Yes",
      TRUE ~ "No"
    )
  )
```
```{r}
# Function to calculate attrition cost
calculate_attrition_cost <- function(
  
  # Employee
  n                    = 1,
  salary               = 80000,
  
  # Direct Costs
  separation_cost      = 500,
  vacancy_cost         = 10000,
  acquisition_cost     = 4900,
  placement_cost       = 3500,
  
  # Productivity Costs
  net_revenue_per_employee = 250000,
  workdays_per_year        = 240,
  workdays_position_open   = 40,
  workdays_onboarding      = 60,
  onboarding_efficiency    = 0.50
  
) 
{
  # Direct Costs
  direct_cost <- sum(separation_cost, vacancy_cost, acquisition_cost, placement_cost)
  
  # Lost Productivity Costs
  productivity_cost <- net_revenue_per_employee / workdays_per_year *
    (workdays_position_open + workdays_onboarding * onboarding_efficiency)
  
  # Savings of Salary & Benefits (Cost Reduction)
  salary_benefit_reduction <- salary / workdays_per_year * workdays_position_open
  
  # Estimated Turnover Per Employee
  cost_per_employee <- direct_cost + productivity_cost - salary_benefit_reduction
  
  # Total Cost of Employee Turnover
  total_cost <- n * cost_per_employee
  
  return(total_cost)
  
}

calculate_attrition_cost()
calculate_attrition_cost(200)
```
```{r}
# Function creation
dept_job_role_tbl %>%
  
  # Block 1
  group_by(Department, JobRole, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  
  # Block 2
  group_by(Department, JobRole) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  
  # Block 3
  filter(Attrition %in% "Yes") %>%
  arrange(desc(pct)) %>%
  mutate(
    above_industry_avg = case_when(
      pct > 0.088 ~ "Yes",
      TRUE ~ "No"
    )
  ) %>%
  
  # Block 4. Set salary to 80000 for now
  mutate(
    cost_of_attrition = calculate_attrition_cost(n = n, salary = 80000)
  )
```
```{r}
dept_job_role_tbl %>% count(Department, JobRole, Attrition)
# Function to convert counts to percentages. 
count_to_pct <- function(data, ..., col = n) 
{
  # Capture the dots
  grouping_vars_expr <- quos(...)
  col_expr <- enquo(col)
  ret <- data %>%
    group_by(!!! grouping_vars_expr) %>%
    mutate(pct = (!! col_expr) / sum(!! col_expr)) %>%
    ungroup()
  return(ret)
}

dept_job_role_tbl %>%
  count(JobRole, Attrition) %>%
  count_to_pct(JobRole)
  
dept_job_role_tbl %>%
  count(Department, JobRole, Attrition) %>%
  count_to_pct(Department, JobRole) 
```
```{r}
dept_job_role_tbl %>%
  group_by(Department, JobRole, Attrition) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(Department, JobRole) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  filter(Attrition %in% "Yes") %>%
  arrange(desc(pct)) %>%
  mutate(
    above_industry_avg = case_when(
      pct > 0.088 ~ "Yes",
      TRUE ~ "No"
    )
  ) %>%
  mutate(
    cost_of_attrition = calculate_attrition_cost(n = n, salary = 80000)
  )
```
```{r}
assess_attrition <- function(data, attrition_col, attrition_value, baseline_pct) 
{
  attrition_col_expr <- enquo(attrition_col)
  data %>%
    # Use parenthesis () to give tidy eval evaluation priority
    filter((!! attrition_col_expr) %in% attrition_value) %>%
    arrange(desc(pct)) %>%
    mutate(
      # Function inputs in numeric format (e.g. baseline_pct = 0.088 don't require tidy eval)
      above_industry_avg = case_when(
        pct > baseline_pct ~ "Yes",
        TRUE ~ "No"
      )
    )
}
 
# Visualization
dept_job_role_tbl %>%
  count(Department, JobRole, Attrition) %>%
  count_to_pct(Department, JobRole) %>%
  assess_attrition(Attrition, attrition_value = "Yes", baseline_pct = 0.088) %>%
  mutate(
    cost_of_attrition = calculate_attrition_cost(n = n, salary = 80000)
  ) %>%
  # Data Manipulation
  mutate(name = str_c(Department, JobRole, sep = ": ") %>% as_factor()) %>%
  # Check levels
  mutate(name      = fct_reorder(name, cost_of_attrition)) %>%
  mutate(cost_text = str_c("$", format(cost_of_attrition / 1e6, digits = 2),
                           "M", sep = "")) %>%
  
  # Plotting
  ggplot(aes(cost_of_attrition, y = name)) +
  geom_segment(aes(xend = 0, yend = name),    color = "#2dc6d6") +
  geom_point(  aes(size = cost_of_attrition), color = "#2dc6d6") +
  scale_x_continuous(labels = scales::dollar) +
  geom_label(aes(label = cost_text, size = cost_of_attrition),
             hjust = "inward", color = "#2dc6d6") +
  scale_size(range = c(3, 5)) +
  labs(title = "Estimated cost of Attrition: By Dept and Job Role",
       y = "",
       x = "Cost of attrition") +
  theme(legend.position = "none")
```
```{r}
# Descriptive Features
employee_attrition_tbl %>% select(Age, DistanceFromHome, Gender, MaritalStatus, NumCompaniesWorked, Over18)
```
```{r}
# Employment Features
employee_attrition_tbl %>% select(Department, EmployeeCount, EmployeeNumber, JobInvolvement, JobLevel, JobRole, JobSatisfaction)
```
```{r}
# Compensation Features
employee_attrition_tbl %>% select(DailyRate, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, StockOptionLevel)
```
```{r}
# Survery Results
employee_attrition_tbl %>% select(EnvironmentSatisfaction, JobSatisfaction, RelationshipSatisfaction, WorkLifeBalance)
```
```{r}
# Performance Data
employee_attrition_tbl %>% select(JobInvolvement, PerformanceRating)
```
```{r}
# Work-Life Features
employee_attrition_tbl %>% select(BusinessTravel, OverTime)
```
```{r}
# Training & Education
employee_attrition_tbl %>% select(Education, EducationField, TrainingTimesLastYear)
```
```{r}
# Time-Based Features
employee_attrition_tbl %>% select(TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)
```
```{r}
library(GGally)
# Data Visualization ----
employee_attrition_tbl %>%
  select(Attrition,  Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
  ggpairs()
```
```{r}
# Explore Features by Category
plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) 
{
  color_expr <- enquo(color)
  if (rlang::quo_is_null(color_expr)) 
  {
    g <- data %>% ggpairs(lower = "blank") 
  } 
  else 
  {
    color_name <- quo_name(color_expr)
    g <- data %>%
      ggpairs(mapping = aes_string(color = color_name), 
              lower = "blank", legend = 1,
              diag = list(continuous = wrap("densityDiag", 
                                            alpha = density_alpha))) +
      theme(legend.position = "bottom")
  }
  return(g)
}
```

```{r}
#   1. Descriptive features: age, gender, marital status 
employee_attrition_tbl %>%
  select(Attrition, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   2. Employment features: department, job role, job level
employee_attrition_tbl %>%
  select(Attrition, contains("employee"), contains("department"), contains("job")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   3. Compensation features: HourlyRate, MonthlyIncome, StockOptionLevel 
employee_attrition_tbl %>%
  select(Attrition, contains("income"), contains("rate"), contains("salary"), contains("stock")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   4. Survey Results: Satisfaction level, Work Life Balance 
employee_attrition_tbl %>%
  select(Attrition, contains("satisfaction"), contains("life")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   5. Performance Data: Job Involvement, Performance Rating
employee_attrition_tbl %>%
  select(Attrition, contains("performance"), contains("involvement")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   6. Work-Life Features 
employee_attrition_tbl %>%
  select(Attrition, contains("overtime"), contains("travel")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   7. Training and Education 
employee_attrition_tbl %>%
  select(Attrition, contains("training"), contains("education")) %>%
  plot_ggpairs(Attrition)
```
```{r}
#   8. Time-Based Features: Years at company, years in current role
employee_attrition_tbl %>%
  select(Attrition, contains("years")) %>%
  plot_ggpairs(Attrition)
```
# Q1: Compensation Features (Monthly income ~ Attrition)
# A1: c

# Q2: Compensation Features (Percent Salary Hike ~ Attrition)
# A2: d

# Q3: Compensation Features (Stock Option Level ~ Attrition)
# A3: a

# Q4: Survey Results (Environment Satisfaction ~ Attrition)
# A4: c

# Q5: Survey Results (Work Life Balance ~ Attrition)
# A5: a

# Q6: Performance Data (Job Involvement ~ Attrition)
# A6: c

# Q7: Work-Life Features (Over Time ~ Attrition)
# A7: a

# Q8: Training and Education (Training Times Last Year ~ Attrition)
# A8: c

# Q9: Time-Based Features (Years At Company ~ Attrition)
# A9: a

# Q10: Time-Based Features (Years Since Last Promotion ~ Attrition)
# A10: a